3D Environment Generation Using Conditional Adversarial Networks

### 2019-12-10 writing/paper/paper.tex
Add related and future work

@@

	This project explores the new paradigm of 3 dimensional procedurally generated content via machine learning.

@@

	Procedural content generation is an algorithmic and stochastic process used in art, games, music and film. As a paradigm for content generation, it is used to increase replayability, variability, flexibility and reduce production cost, effort storage space \cite{summerville2017procedural}.

@@

	Neural networks allow for a powerful model of learning patterns from complex datasets. Generative adversarial networks work by synthesizing novel data according to patterns it learns in the training set. We evaluate two models of conditional generative adversarial networks to achieve image to image translation.

	We are going to evaluate two implementations of a specialized form of generative adversarial networks designed to perform conditional image to image translation. The two models we are using are called Pix2Pix and CycleGAN.

	Both Pix2Pix and CycleGAN are based on the deep convolutional generative adversarial network architecture proposd by Radford et al. in 2015 \cite{radford2015unsupervised}.

	The objective function of a conditional GAN can be expressed as:

@@

	CycleGAN introduces a cycle consistency loss to their objective function. The idea is that for each image in the $x$ domain, the image translation cycle should be able to bring the $x$ back to the original image, $x \rightarrow G(x) \rightarrow F(G(x)) \approx x$.

@@

	Nvidia GauGAN.

@@

	In this section we go over all of the pieces of work that went in to this project.

@@

	We also find that the model does not perform very well on usermaps with rivers. We found that rivers tend to produce noisy results and sometimes the rivers aren't defined enough to be noticeable in the results. I think this can be solved with more training data including rivers. Due to the nature of how the training data was gathered, it's likely that we didn't get enough examples, and due to the subjective nature of myself reading the earth data and drawing the usermaps myself, I could have read them wrong and not painted them as detailed as thy needed to be.

@@

@@

		\end{tabular}
	\end{table*}

@@

>>>>

	This approach offers a promising avenue to co-creative design. Humans working together with machine learning programs could allow for more creative results with reduced cost and effort for the designer. The flexibility of being able to tweak and tune the training set also allows for the possibility of adapting this approach to content of many different kinds, not just 3d environments.

@@

@@

	Procedural content generation is an algorithmic and stochastic process used in art, games, music and film. As a paradigm for content generation, it is used to increase replayability, variability, flexibility and reduce production cost, effort and storage space \cite{summerville2017procedural}. These algorithms are crafted by hand and they require a lot of tuning by a developer to ensure that the random perturbations result in usable content.

	Procedral content generation has been used in production applications since 1980 with the dungeon crawler Rogue demonstrating its effectiveness. \cite{wiki:procedural} Procedural generation has been a staple in gaming since and today it continues to be featured in some of the top games.

	The addition of machine learning to procedural generation is a new topic in research. Most projects are using 2 dimensional platformers, but real time strategy and 3 dimensional environments are also being explored. This field of research has its own challenges, particularly the comparability of results, but efforts are beng made to improve benchmarks and datasets \cite{summerville2017procedural}.

@@

	Neural networks allow for a powerful model of learning patterns from complex datasets. Generative adversarial networks work by synthesizing novel data according to patterns it learns in the training set. They work by training two neural networks in competition with each other, a generator and a discriminator. The generator network generates new data instances and the discriminator classifies the results of the generator according to how authentic the generated images are given some training distribution. These two networks train in tandem, and over time, the generator gets good at producing novel images that look like they belong to the training set.

	We are evaluate two implementations of a specialized form of generative adversarial networks designed to perform conditional image to image translation. The two models we are using are called Pix2Pix and CycleGAN.

	Both Pix2Pix and CycleGAN are conditonal networks based on the deep convolutional generative adversarial network architecture proposed by Radford et al. in 2015 \cite{radford2015unsupervised}.

	The objective function of a conditional generative adversarial networks can be expressed as:

@@

	Where pix2pix trains one GAN (one generator and one discriminator), CycleGAN trains two, (Generator A and B and Discriminator A and B). learning the mapping between the two GANs is the trick to getting unpaired image translation to work. CycleGAN introduces a cycle consistency loss. The idea is that for each image in the $x$ domain, the image translation cycle should be able to bring the $x$ back to the original image, $x \rightarrow G(x) \rightarrow F(G(x)) \approx x$.

@@

	\subsubsection{Nvidia's GauGAN}

	GauGAN was the inspiration behind the semantic user map. This research demonstrated an application that took a 2 dimensional semantic image drawn by a user, and translated it into a photorealistic image. This project also used Pix2Pix as a base for their network architecture, but they also introduce a novel normalization layer called spatially-adaptive normalization. The idea is that normal layer normalization tends to wash out semantic meaning from an image. By considering the spatial properties, you can keep the semantic information in tact while still performing normalization. \cite{park2019semantic}	. Nvidia hosts an excellent demo of the project at http://nvidia-research-mingyuliu.com/gaugan.

	\subsubsection{A Step Toward Procedural Terrain Generation with GANs}

	This paper had a very similar idea to this project. They use a standard generative adversarial network to synthesize terrain heightmaps with a model trained on Earth data. They use NASA's visible earth project for their training image data at a resolution of 1km per pixel. This project appears to be quite introductory and they lay out a good overview of how a standard GAN could be implemented to generate environments. We expand on their work by introducing the conditional GAN and integrating co-creative design by including a translation from a user defined segmentation map. \cite{beckham2017step}

@@

@@

	We also find that the model does not perform very well on usermaps with rivers. We found that rivers tend to produce noisy results and sometimes the rivers aren't defined enough to be noticeable in the results. I think this can be solved with more training data including rivers. Due to the nature of how the training data was gathered, it's likely that we didn't get enough examples, and due to the subjective nature of myself reading the earth data and drawing the usermaps myself, I could have read them wrong and not painted them as detailed as they needed to be.

@@

	\subsection{User Preference Survey}

	We also created a small user survey to measure user preference between the 4 trained models. The survey was conducted through google forms and sent to the students in CSCI 567 (3D User Interfaces) through the piazza message board. I took a sample of 5 user maps, and for each one, I produced the ouput for each model and listed those as multiple choice answers.

@@

		\end{tabular}

		\caption{Test images and the outputs from our trained models}
	\end{table*}

@@

	Expanding this idea to include texture maps could be an interesting direction to go. Using this same method, it should be possible to translate a given heightmap into a texturemap. This could allow the designer to have control over generated styles like barren desert scenes vs dark horror scenes vs fantastical colorful scenes. Gathering training data of texturemaps would likely be the bottleneck.

	Modifying the network to perform real time generation would be an interesting addition. This would give the user a more interactive experience in designing the enviroments. Current latency with this implementation is not fast enough so new network architectures or approaches would be needed.

	Tunable paramaters would also be an interesting addition. Giving the user even more control over the results would lead to a better experience. I can imagine parameters like intensity, noise, smoothness could rusult in a lot of usable outcomes.

	Expanding the training set with different means. Needing to draw all the usermaps for the training set is time consuming and error/bias prone. Coming up with a programmatic way to create the usermap training set could have a large impact on the quality of the results.
